### 2012-02-16 2.1.1.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.1.1: Data sources

The English CELEX database \citep{CELEX},

\citet{Aronoff1976}

\citet{Harley2009}

\citet{Taft1975} ?
\citet{Taft1981} ?
\citet{Taft2004a} ?

\citet{Halle1962}