### 2012-02-17 3.1.1.tex

% 3.1.1: Alternation vs. phonotactic

\subsubsection{Roundness harmony}

\subsubsection{Labial attraction}

\citet{Zuraw2000}
\citet{Clements1982}
\citet{Inkelas1997}