### 2012-02-17 3.1.2.tex

% 3.1.2: Lexical statistics

This is essentially the hypothesis pursued in the previous chapter.

\citet{TELL}
\citet{Inkelas2001}

%We might, for instance, consider methods based not on statistical \citep{Yang2005,Gorman2011b}.