### 2012-02-17 3.2.2.tex

% 3.2.2: Wordlikeness results

\citet{Goodman1954}

\ex $\displaystyle \gamma = \frac{C - D}{C + D}$ \xe

\subsubsection{Roundness harmony}

($\gamma = 0.597$, $p = 5.2$\e{-29})

Roundness harmony wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
\midrule
pemez & 30                   & pemaz & 2 \\
teriz & 28                   & terız & 3 \\
tokaz & 26                   & tokez & 6 \\
tipez & 24                   & tipaz & 8 \\
terüz & 19                   & teruz & 13 \\
\bottomrule
\end{tabular}

\subsubsection{Labial harmony}

Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{aPu} & \multicolumn{2}{l}{aPı} \\
\midrule
tafuz & 21              & tafız & 11 \\
pamuz & 17              & pamız & 15 \\
tapuz & 17              & tapız & 15 \\
mavuz & 16              & mavız & 16 \\
tabuz & 16              & tabız & 16 \\
\bottomrule
\end{tabular}

($\gamma = 0.087$, $p = 0.101$)