### 2012-02-17 4.1.2.tex

% 4.1.2: The ontology of wordlikeness

\subsubsection{Gross phonotactic violation}

position in \citet{SPE}

\citet{Shibatani1972}
\citet{Hooper1973}

peripheral \citep[in the sense of][81]{LGB}

As an example of this applied ot phonotactics, 
\citet{Borowsky1989} assumes that there are peripheral morph-internal syllable contact clusters.

\subsubsection{Phonological statistics}

\citet{SPE}

This statement is made very strongly by \citet{Coleman1996}

\subsubsection{Lexical similarity}

\citet{LSLT}

\citet{Greenberg1964}
\citet{Ohala1986b}

A recent, very articulated model is presented by \citet{Bailey2001}.

% simple model
This is simply because more complex models fail to

% similarity vs. rules
% as a final note, \citet{Hahn1998}