### 2012-02-17 4.2.1.tex

% 4.2.1: Assessing gradience

\subsubsection{Response distributions}

This is shown in Figure \ref{density}.

\begin{figure}
\centering
\includegraphics{density.pdf}
\caption{Density of responses}
\label{density}
\end{figure}

\subsubsection{Prediction distributions}

\citet{Coltheart1977}

\citet{Hayes2008a}
\citet{Albright2009}

% stats thereof
%\citet{Schilling2002}
%\citet{Helguerro1904}
%\citet{Cohen1956}

% EM
%\citet{EM}

% -2 log like ratio test