### 2012-02-17 4.2.2.tex

% 4.2.2: Assessing models

\subsubsection{Basic correlations}
\subsubsection{Intramodal correlations}
\subsubsection{Residual correlations}