### 2012-03-01 3.1.tex

% 3.1: Wordlikeness predictions

\citet{Zimmer1969}

I assume without discussion the following binary feature specifications for the vowel phonemes of Turkish.

\ex Turkish vowel features: \vspace{6pt} \\
\begin{tabular}{c | c c c c}
                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
\end{tabular}
\xe

%\citep[][289]{Anderson1975}
%\citep{Clements1982}
%\citet{Inkelas1997}
%\citet{Harrison2001}

\subsection{Backness harmony}  \input{3.1.1}
\subsection{Labial attraction} \input{3.1.2}