### 2012-03-01 4.3.tex

% 4.3: Conclusion

It is not the case that this evidence is inconsistent with the hypothesis that there are additional degrees of wordlikeness: it is just a demonstration that these data to not require it.