### 2012-03-01 5.tex

\chapter{Inflectional gaps in generative grammar}

\section{A brief history of inflectional gaps} \input{5.1}
\section{The null hypothesis}                  \input{5.2}
\section{Outline of Part II}                   \input{5.3}