### 2012-03-01 7.1.1.tex

\subsection{Proposal}

\citet{OT}
\citet{Orgun1999}

\subsubsection{Inviolable constraints in a rule-based phonology}

\citet{Vaux2008}
(28 ``violability'')

such as developed by Andrea \citet{Calabrese1995,Calabrese2005}.

citing the OCP as an example...it's not clear wtf he's talking about
\citet{Goldsmith1976}

\citet{Nevins2003} % shm-gaps
\citet{Rice2007} % is being produced but avoided?

\citet[][9f.]{Baerman2010b}
\citet[][437]{Heath2005}

\citet{OT}
\citet{Orgun1999}

\subsubsection{MParse}

\citet{OT}
\citet{Raffelsiefen1999}
\citet{Raffelsiefen2004}

\citet{Wolf2009}

\subsubsection{Control}

\citet{Orgun1999}
\citet{Orgun2009}