### 2012-03-01 7.1.2.tex

\subsection{Evaluation}

% 7.1.3.1: Turkish and Turkish'

\citet{Orgun1999}
\citet{Orgun2009}

\citet{Ito1989b}

\subsubsection{Kinande reduplication}

``hurriedness or repetitive action \citep[][85; henceforth, MH]{Mutaka1990}.

\citet{Orgun1999} provide an analysis of a similar gap in Tiene, but it appears that the judgements they report are inferred, as \citet{Ellington1977}, the source for the generalization they report, provides no relevant judgements or examples, and his discussion suggests the gap is morphological in nature.
Since I have been unable to find any other published material on Tiene reduplication, Kinande, which is in comparison well documented, is presented here in its place.
But I predict that any close study of Tiene reduplication will reveal similar lexical exceptions.

\footnote{Thanks to Toni Cook for bringing this example to my attention.}

The material making up the so-called ``pre-stem'' will not be a major object of concern here: it consists of the nominal augment vowel (\textsc{Aug}), a verbal prefix (\textsc{Pre}), and where appropriate, the object marker (\textsc{OM}). Valence-chagning suffixes, known as ``extensions'', may appear to the right (in the below example, the applicative \textsc{Appl} and the reciprocal \textsc{Recp}), and the final \textsc{FV}, which indicates aspect.

\ex Linear structure of the Kinande infinitive: \\
\begin{tabular}{l l l l l l l}
e- & -rí- & -bí & tum & -ir & -an & -a \\
\{\textsc{Aug} & \textsc{Pre} & \texsc{OM}\} & \{send & \textsc{Appl} & \textsc{FV}\} \\
\multicolumn{7}{l}{``to send them to each other''}
\xe

bisyllabic
\citet{Downing2000}

\citet{Mutaka2008}

\subsubsection{Swedish indefinite adjectives}

\citet{Lowenadler2010a}
\citet{Lowenadler2010b}

\citet[][ chapter 5]{Lofstedt2010}

\citet[][7]{Buchanan2007}

\citet{Linell1973}
\citet{Sunden1869}
\citet{Hylen1920}
\citet{Cederschiold1912}
\citet{Brunnmark1826}
\citet{Eliasson1973}
\citet{Sigurd1965}
\citet{Teleman1999}
\citet{Kiefer1970}
\citet{Hellberg1972}
\citet{Iverson1981}
\citet{Pettersson1990}

\subsubsection{Tagalog infixation}

\citet{OT}

\citet{Orgun1999}
\citet{Orgun2009}

\citet{Ross1995}

\citet{Baerman2010b}