### 2012-03-04 3.1.tex
Writing on chapter 3

\section{Turkish vowel sequence structure constraints}

Below, I detail two sequence structure constraints on root vowels in Turkish proposed by \citet{Lees1966a,Lees1966b}, relating these generalizations to Turkish alternation phonology and various sources of external evidence. A standard binary feature specification for the Turkish vowel system, given below, is assumed throughout.

\ex Turkish vowel features: \vspace{6pt} \\
\begin{tabular}{c | c c c c}
                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
\end{tabular} \xe

\input{3.1.1}
\input{3.1.2}

%\citep[][289]{Anderson1975}
%\citep{Clements1982}
%\citet{Inkelas1997}
%\citet{Harrison2001}