### 2012-04-18 2.2.2.tex

\subsection{Assessing models}

\subsubsection{Basic correlations}
\subsubsection{Intramodal correlations}
\subsubsection{Residual correlations}