### 2012-04-18 2.3.tex

\section{Conclusions}

It is not the case that this evidence is inconsistent with the hypothesis that there are additional degrees of wordlikeness: it is just a demonstration that these data to not require it.