### 2012-04-18 3.1.tex

\section{Turkish vowel sequence structure constraints}

\citet{Lees1966a,Lees1966b} proposes three sequence structure constraints on Turkish vowels, and these constraints have been discussed extensively in the subsequent literature. Below, these constraints are formalized and related to Turkish alternation phonology and various sources of external evidence. A traditional binary feature specification for the Turkish vowel system, given below, is assumed without discussion.

\begin{example}
Turkish vowel features:

\vspace{0.5\baselineskip}
\begin{tabular}{c | c c c c}
                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
\end{tabular}
\end{example}

\input{3.1.1}
\input{3.1.2}
\input{3.1.3}