### 2012-04-18 4.1.1.tex

\subsection{Data sources}

The English CELEX database \citep{CELEX},

\citet{Aronoff1976}

\citet{Harley2009}

\citet{Taft1975} ?
\citet{Taft1981} ?
\citet{Taft2004a} ?

\citet{Halle1962}

% 2.1.4: k

\subsubsection{Corpus counts}

Under the phonological analyses described in the next two sections, the 6,876 simplex words in CELEX contain 21 unique medial codas and 40 unique medial onsets. Of these 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, for a 18.8\% saturation rate.