### 2012-04-26 3.1.tex

\section{Turkish vowel sequence structure constraints}

\citet{Lees1966a,Lees1966b} proposes three constraints on Turkish vowel sequences 
which have been the attention of many subsequent studies on vowel harmony and on sequence structure constraints. In this section, these constraints are formalized, and where possible, related to phonological alternations in Turkish and to various sources of external evidence. The following feature specification for the Turkish vowels is assumed throughout.

\begin{example}
Turkish vowel features:

\vspace{0.5\baselineskip}
\begin{tabular}{c | c c c c}
                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
\buf[$+$\textsc{High}] & {i} & {ü} [y] & {ı} [ɯ] & {u} \\
\buf[$-$\textsc{High}] & {e} & {ö} [ø] & {a} [ɑ] & {o} \\
\end{tabular}
\end{example}

\input{3.1.1}
\input{3.1.2}
\input{3.1.3}