### 2012-05-07 3.1.2.tex

\subsection{Roundness Harmony}

Roundness harmony is a closely related process with a somewhat narrower scope.

\subsubsection{Phonological description}

Roundness harmony is quite similar to \textsc{Backness Harmony}, with the additional stipulation that the target must be [$+$\textsc{High}].

\begin{example}
\textsc{Roundness Harmony}:

\xymatrix@R=24pt@C=24pt{
\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{--}[drr] &             & \txt{[$+$\textsc{High}]} \\
\txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
}
\end{example}

This rule, in concert with \textsc{Backness Harmony}, accounts for the forms of the dative singular (dat.sg.) and genitive singular (gen.sg.), among other suffixes.

\begin{example}
Turkish nominal suffix allomorphy:

%\vspace{0.5\baselineskip}
\begin{tabular}{l l l l l l l}
   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
a. & ip             & ipi            & ipin           & `rope' & \citep[][216]{Clements1982} \\
   & el             & eli            & elin           & `hand'    \\
   & kız            & kızı           & kızın          & `girl'    \\
   & sap            & sapı           & sapın          & `stalk'   \\
   & yüz            & yüzü           & yüzün          & `face'    \\
   & köy            & köyü           & köyün          & `village' \\
   & pul            & pulu           & pulun          & `stamp'   \\
   & son            & sonu           & sonun          & `end'     \\
b. & boğaz          & boğazı         & boğazın        & `throat'  & (TELL) \\
   & pelür          & pelürü         & pelürün        & `tissue paper' \\
   & döviz          & dövizi         & dövizin        & `currency' \\
   & yamuk          & yamuğu         & yamuğun        & `trapezoid' \\
   & ümit           & ümiti          & ümitin         & `hope'     \\
\end{tabular}
\end{example}

\noindent As was the case for \textsc{Backness Harmony}, disharmonic roots are found, and the final vowel of disharmonic roots triggesr suffix harmony.

\subsubsection{External evidence}

Though few studies have directly sought external evidence for speaker's awareness of root \textsc{Roundness Harmony}, two of the findings reviewed in \S\ref{backharmexternal} above speak in support of this claim. First, the epenthetic vowel in non-native word-initial onsets
like \emph{trablus} \alt{} \emph{tırablus} `Tripoli' conform also to \textsc{Roundness Harmony}.
%\citealp[][247]{Clements1982}). 
Second, \textsc{Roundness Harmony} participates in reharmonization in \citeauthor{Harrison2001}'s language game: the reduplicated \emph{bütün-batın} exhibits replacement of the second \emph{ü}, a [$+$\textsc{Round}] vowel, with [$-$\textsc{Round}] \emph{ı}.