### 2012-05-07 ack.tex

So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy. 
More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, 
who did some much to make this work, and me as a linguist,
into what it is now.
Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are to be thanked for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
And though his name appears nowhere else on this document, it is hard to imagine this work could exist without Bill Labov.

My journey in linguistics begins with the mentoring of Jennifer Cole.
Steve Anderson graciously agreed to serve on my committee.
Morris Halle taught me to speak clearly and never apologize.

While so many of my fellow students at Penn have a hand in this document, 
I would like to especially thank my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody.

%Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document.
%Maryam Bakht kept me centered.

My gratitude also goes out to the administrative assistants at 3401 Walnut and 255 S. 36th, Lauris Olson at the Van Pelt Library, Katherine McEldoon for a helpful review of numerical cognition research, and Carly Fisher for her unflagging support and friendship.