### 2012-06-30 ack.tex
Feverish writing on chapters 2-4

So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy. More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, who has done so much for me as a linguist. Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are to be thanked for creating tan environment receptive to this admittedly contentious thesis. And though his name appears nowhere else in this document, it is hard to imagine this dissertation would exist without Bill Labov.

While so many of my fellow students at Penn have a hand in this document, I would like to especially thank Toni Cook and Laurel MacKenzie for all their support over the years. I also must thank the administrative assistants at 3401 Walnut and 255 S. 36th, Lauris Olson at the Van Pelt Library, Katherine McEldoon for a helpful review of numerical cognition research, and Carly Fisher for ten years of her unflagging friendship.

I have many debts beyond Penn. My journey in linguistics began with the careful mentoring of Jennifer Cole and Richard Sproat. Steve Anderson graciously agreed to serve on my committee. Morris Halle taught me to speak clearly, and to never apologize.

%Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document.
%Maryam Bakht kept me centered.