### 2012-10-10 ack.tex

So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy. More than anyone else, I owe a debt of gratitude to Charles Yang, who has done (and put up with) so much for me. I am grateful to my committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment for quantitative and theoretical approaches to meet. And though his name does not appear again in this document, so much here has been inspired by Bill Labov.

The faculty, staff, and students of the Institute For Research in Cognitive Science have provided me with not merely an office, but a home, complete with dated early-90s decor. Thanks to my officemates, Chandan Narayan, John Trueswell, Laurel Sweeney and the rest of the staff there, as well as Amy Forsyth back at Williams Hall.

So many of my fellow students at Penn have had a hand in this document, none so much as Toni Cook and Laurel MaKenzie, who have supported me so much over the past few years. Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun. Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar.

The ideas I first encountered under the tutelage of Jennifer Cole and Richard Sproat at the University of Illinois continue to guide my research. Insofar as I have acknowledged historical precursors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee. Morris Halle taught me to speak clearly and to never apologize.