### 2012-11-05 ack.tex

So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy.

The ideas I first encountered under the tutelage of Jennifer Cole and Richard Sproat at the University of Illinois continue to guide my research.

The faculty, staff, and students of the Institute For Research in Cognitive Science provided me with not merely an office, but something close to a home. Thanks to my officemates, Chandan Narayan, John Trueswell, Laurel Sweeney and the rest of the staff there, as well as Amy Forsyth back at Williams Hall.

So many of my fellow students at Penn have had a hand in this document, none so much as Toni Cook and Laurel MaKenzie, who have supported me so much over the past few years. Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun. Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar. John Murdock fearlessly proofread everything, twice.

I am grateful to my committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment for quantitative and theoretical approaches to meet.

Insofar as I have acknowledged my historical precursors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee.

More than anyone else, I owe a debt of gratitude to Charles Yang: I don't know how he puts up with me.

I would like to conclude by thanking three people who were \emph{not} involved in the preparation of this dissertation, but without which it could never exist: Lila Gleitman, Morris Halle, and Bill Labov.