### 2013-01-28 ack.tex

While many of my fellow students at Penn have had a hand in this document, none so much as my cohort: Toni Cook and Laurel MaKenzie. Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun. Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar. John Murdock and Hilary Prichard proofread the entire document numerous times.

The faculty, staff, and students of the Institute For Research in Cognitive Science provided me with not merely an office, but something close to a home. Thanks to my officemates, Jessica Marcus, Chandan Narayan, John Trueswell, and especially Laurel Sweeney.

The ideas I first encountered as an undergraduate at the University of Illinois under the tutelage of Jennifer Cole and Richard Sproat, continue to guide my research. I am grateful to the members of the committee, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment where quantitative and theoretical approaches can meet. Insofar as I have acknowledged and understood the arguments of my predecessors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee and to show me the wonders of New Haven. 
I do not expect to ever again have such a fearless advocate in the incomporable Charles Yang; the scientist he is is the one I strive to be.

While they were not involved in the preparation of this dissertation, it is hard to imagine that it would exist without Lila Gleitman, Morris Halle, and Bill Labov.